Paper Wireframe

### 2018-07-31 paper_wireframe_knit.Rmd
Fixes so code works with expanded analyses

@@

dat2.2<-dat2.1 %>% group_by(year, ptt, SO_startDOY) %>%
  summarise_all(mean, na.rm=T)

@@

So what do these early leaving birds do in their runup to departing Cote d'Ivoire that is different to other birds?

@@

soversWA2<-left_join(soversWA1, dat2.2[,c(1,2, 3,7)], by=c('ptt', 'year', 'SO_startDOY'))

@@

dat_ls

@@

  addCircleMarkers(data=d2, popup= paste("ptt:", d2$ptt, "<br>", "year:", d2$year,"<br>", "NDVIanom:", d2$aquaANOM), color=~ifelse(aquaANOM>0, 'green', 'red'), stroke=F, radius=5, group='Early departure') %>%

    addCircleMarkers(data=d3, popup= paste("ptt:", d3$ptt, "<br>", "year:", d3$year,"<br>", "NDVIanom:", d3$aquaANOM), color=~ifelse(aquaANOM>0, 'green', 'red'), stroke=F, radius=5, group='Late departure') %>%

@@

dat2<- env %>% group_by(year, ptt, SO_startDOY, variable) %>%
  summarise(country=last(country), SO_endDOY=first(SO_endDOY),
            cuck_pres=first(cuck_pres), precip=mean(value, na.rm=T),
            NDVI= mean(value2, na.rm=T), aquaNDVI=mean(aquaNDVI, na.rm=T),
            aquaANOM=mean(aquaANOM, na.rm=T),terrNDVI=mean(terrNDVI, na.rm=T), terrANOM=mean(terrANOM, na.rm=T), tamRAIN=mean(tamRAIN, na.rm=T), tamRAINanom=mean(tamRAINanom, na.rm=T) )

dat2<- dat2[- which(dat2$country %in% c('Angola', 'Central African Republic','Congo', 'Democratic Republic of the Congo', 'Gabon',  'Cameroon', 'Burkina Faso', 'Liberia')),]

# relevel to central well sampled year
dat2$country<-factor(dat2$country, levels=c('Nigeria', 'Togo', 'Ghana', "Cote d'Ivoire", 'Guinea', 'Sierra Leone'))

ggplot()+geom_line(data=dat2[dat2$cumrf>0,],aes(x=variable, y=cumrf, group=paste(ptt, year,SO_startDOY)), size=2, alpha=0.08, colour='blue')+theme_bw()+geom_vline(data=(dat2 %>% group_by(ptt, year, SO_startDOY) %>% summarise_all(first)), aes(xintercept = SO_startDOY), colour='red', alpha=0.5)+
  geom_vline(data=(dat2[dat2$cuck_pres=='Y',] %>% group_by(ptt, year) %>% summarise_all(first)), aes(xintercept = SO_startDOY), colour='black', alpha=0.5)+facet_wrap(~country)+xlab("Day of year")+
  ylab("Cumulative rainfall (mm)")

@@

Shown with mean rainfall per country (blue line) and mean arrival date for first stopovers in West Africa (red line), we can see rains arrive at generally the same time across in Cote d'Ivoire than in Ghana and so are cuckoos?

temp<-dat2[dat2$cuck_pres=='Y',] %>% group_by(ptt, year) %>% summarise_all(first) %>% group_by(country) %>% summarise(MSO_startDOY=mean(SO_startDOY), Mrf=mean(cumrf), Mnd=mean(aquaNDVI))

temp2<-dat2 %>% group_by(country, variable) %>% summarise(Mcumrf=mean(cumrf), rfmin=mean(cumrf)-sd(cumrf), rfmax=mean(cumrf)+sd(cumrf),                                                          MaquaNDVI=mean(aquaNDVI, na.rm=T), ndmin=(mean(aquaNDVI, na.rm=T)-sd(aquaNDVI, na.rm=T)), ndmax=(mean(aquaNDVI, na.rm=T)+sd(aquaNDVI, na.rm=T)))

p1<-ggplot()+
geom_line(data=temp2,aes(x=variable, y=Mcumrf), size=1, colour='blue')+
geom_line(data=temp2,aes(x=variable, y=rfmin), colour='grey', linetype=2)+
geom_line(data=temp2,aes(x=variable, y=rfmax), colour='grey', linetype=2)+  
geom_vline(data=(dat2[dat2$cuck_pres=='Y',] %>% group_by(ptt, year) %>% summarise_all(first)), aes(xintercept = SO_startDOY), colour='black', alpha=0.5)+
geom_vline(data=temp, aes(xintercept = MSO_startDOY), colour='red', size=1.5)+
   facet_grid(country~.)+xlab("Day of year")+
  ylab("Cumulative rainfall (mm)")+theme_bw()

p2<-ggplot()+geom_line(data=temp2,aes(x=variable, y=MaquaNDVI), size=1, colour='dark green')+
geom_line(data=temp2,aes(x=variable, y=ndmin), colour='orange', linetype=2)+
geom_line(data=temp2,aes(x=variable, y=ndmax), colour='orange', linetype=2)+
geom_vline(data=(dat2[dat2$cuck_pres=='Y',] %>% group_by(ptt, year) %>% summarise_all(first)), aes(xintercept = SO_startDOY), colour='black', alpha=0.5)+
geom_vline(data=temp, aes(xintercept = MSO_startDOY), colour='red', size=1.5)+
   facet_grid(country~.)+xlab("Day of year")+
  ylab("NDVI")+theme_bw()

@@

temp<-dat2[dat2$cuck_pres=='Y',] %>% group_by(ptt, year) %>% summarise_all(first)

>>>>

temp<- env %>% group_by(year, ptt, SO_startDOY, variable) %>% summarise(country=first(country))

dat2$country<-temp$country

@@

dat2.2<-dat2.1 %>% group_by(year, ptt, SO_startDOY) %>%
  summarise_if(is.numeric, mean, na.rm=T)

@@

So what do these early leaving birds do in their runup to departing Cote d'Ivoire that is different to other birds? The tracks of early departing birds are grouped with blue lines and late departing with orange lines. Each stopover is coloured either red (indicating below average NDVI) or green (above average NDVI) to see if early birds chain together 'greener stopovers.

@@

soversWA2<-left_join(soversWA1, dat2.2[,c(1,2, 3,9)], by=c('ptt', 'year', 'SO_startDOY'))

@@

#dat_ls

@@

  addCircleMarkers(data=d2, popup= paste("ptt:", d2$ptt, "<br>", "year:", d2$year,"<br>", "NDVIanom:", d2$aquaANOM), color=~ifelse(aquaANOM>0, 'green', 'red'), stroke=F, radius=5, opacity=0.2, group='Early departure') %>%

    addCircleMarkers(data=d3, popup= paste("ptt:", d3$ptt, "<br>", "year:", d3$year,"<br>", "NDVIanom:", d3$aquaANOM), color=~ifelse(aquaANOM>0, 'green', 'red'), stroke=F, radius=5,opacity=0.2, group='Late departure') %>%

@@

temp<- dat2[- which(dat2$country %in% c('Angola', 'Central African Republic','Congo', 'Democratic Republic of the Congo', 'Gabon',  'Cameroon', 'Burkina Faso', 'Liberia')),]

temp$country<-factor(temp$country, levels=c('Nigeria', 'Togo', 'Ghana', "Cote d'Ivoire", 'Guinea', 'Sierra Leone'))

ggplot()+geom_line(data=temp[temp$cumrf>0,],aes(x=variable, y=cumrf, group=paste(ptt, year,SO_startDOY)), size=2, alpha=0.08, colour='blue')+theme_bw()+geom_vline(data=(temp %>% group_by(ptt, year, SO_startDOY) %>% summarise_all(first)), aes(xintercept = SO_startDOY), colour='red', alpha=0.5)+
  geom_vline(data=(temp[temp$cuck_pres==2,] %>% group_by(ptt, year) %>% summarise_all(first)), aes(xintercept = SO_startDOY), colour='black', alpha=0.5)+facet_wrap(~country)+xlab("Day of year")+
  ylab("Cumulative rainfall (mm)")

@@

Shown with mean rainfall (blue line) and NDVI (green line) per country and mean arrival date for first stopovers in West Africa (red line), we can see rains arrive at generally the same time in Cote d'Ivoire and Ghana and so do cuckoos

```{r, warning=F, message=F, echo=F, fig.height=12, fig.width=6}

temp<- dat2[- which(dat2$country %in% c('Angola', 'Central African Republic','Congo', 'Democratic Republic of the Congo', 'Gabon',  'Cameroon', 'Burkina Faso', 'Liberia')),]

temp$country<-factor(temp$country, levels=c('Nigeria', 'Togo', 'Ghana', "Cote d'Ivoire", 'Guinea', 'Sierra Leone'))

temp2<-temp[temp$cuck_pres==2,] %>% group_by(ptt, year) %>% summarise_all(first) %>% group_by(country) %>% summarise(MSO_startDOY=mean(SO_startDOY), Mrf=mean(cumrf), Mnd=mean(aquaNDVI))

temp3<-temp %>% group_by(country, variable) %>% summarise(Mcumrf=mean(cumrf), rfmin=mean(cumrf)-sd(cumrf), rfmax=mean(cumrf)+sd(cumrf),                                                          MaquaNDVI=mean(aquaNDVI, na.rm=T), ndmin=(mean(aquaNDVI, na.rm=T)-sd(aquaNDVI, na.rm=T)), ndmax=(mean(aquaNDVI, na.rm=T)+sd(aquaNDVI, na.rm=T)))

p1<-ggplot()+
geom_line(data=temp3,aes(x=variable, y=Mcumrf), size=1, colour='blue')+
geom_line(data=temp3,aes(x=variable, y=rfmin), colour='grey', linetype=2)+
geom_line(data=temp3,aes(x=variable, y=rfmax), colour='grey', linetype=2)+  
geom_vline(data=(temp[temp$cuck_pres==2,] %>% group_by(ptt, year) %>% summarise_all(first)), aes(xintercept = SO_startDOY), colour='black', alpha=0.5)+
geom_vline(data=temp2, aes(xintercept = MSO_startDOY), colour='red', size=1)+
   facet_grid(country~.)+xlab("Day of year")+
  ylab("Cumulative rainfall (mm)")+theme_bw()

p2<-ggplot()+geom_line(data=temp3,aes(x=variable, y=MaquaNDVI), size=1, colour='dark green')+
geom_line(data=temp3,aes(x=variable, y=ndmin), colour='orange', linetype=2)+
geom_line(data=temp3,aes(x=variable, y=ndmax), colour='orange', linetype=2)+
geom_vline(data=(temp[temp$cuck_pres==2,] %>% group_by(ptt, year) %>% summarise_all(first)), aes(xintercept = SO_startDOY), colour='black', alpha=0.5)+
geom_vline(data=temp2, aes(xintercept = MSO_startDOY), colour='red', size=1)+
   facet_grid(country~.)+xlab("Day of year")+
  ylab("NDVI")+theme_bw()

@@

temp<-dat2[dat2$cuck_pres==2,] %>% group_by(ptt, year) %>% summarise_all(first)
